Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-12-27 manuscript/prb-submitted/paper.tex
reply added

%\subsection{Parameters $D_2$ and $D_4$ of the JT ion/atom}
The terms in square brackets on the right of Eqs.~\ref{eq:JTt2g},\ref{eq:JTeg}
contain $D_2$ and $D_4$. 
$D_4$ is proportional to the crystal field splitting $\Delta$ 
($D_4=\frac{3a}{160q_o} \Delta$) and 
we can estimate its value from realistic values of $\Delta$ for a given octahedron of size $a$ and ion charge $q_o$. 
Assuming $a=7.0$ bohr and $q_o=2e$ ($e$ being the elementary charge),
$D_4$ ranges from $1 \times 10^{-3}$ to  $7 \times 10^{-3}$
for $\Delta \simeq 0.5-3$~eV.
Typically, ${D_{2}\approx \sqrt{D_{4}}}$,
so $D_{2}/D_{4} \approx 1/\sqrt{D_{4}}$
and it roughly ranges between $29$ and $12$ for $D_4$ in the above range.
This means that, for realistic systems, 
we should always have $D_{2}/D_{4} > 100/27, 25/9 \approx 4$
so 
$D_{2}$ terms in 
Eqs.~\ref{eq:JTt2g},\ref{eq:JTeg}
dominate and determine the sign and size of the splittings.

@@

In a cubic perovskite ABO$_3$,
each transition metal atom/ion B
is surrounded by six O$^{2-}$, each at a distance $a/2$,
 making an octahedron of size $a$ around it.
The potential of these ions $V_{O}(\vec r)= V_{oct}(\vec r)$
splits the degenerate d-orbitals
by an amount ${\Delta_o=160q_oD_{4}/3a}$
 into $t_{2g}$ and $e_g$.
However, this is not the only crystal field potential at the B site.
The coordination cages of next nearest neighbours, 
eight A atoms/cations all at the same distance $\sqrt{3}a/2$
make a cube around the B ion,
which also produces an octahedral field, given by 
\begin{align}
V_{A}(\vec r) = \frac{8}{81 \sqrt{3}}\frac{q_A}{q_o} V_{O}(\vec r).
\end{align}
Furthermore,
the nearest neighbouring B ions, six of them each at a distance $a$,
make an octahedron, with a potential 
\begin{align}
V_{B}(\vec r) = -\frac{1}{32}\frac{q_B}{q_o}V_{O}(\vec r).
\end{align}
It is interesting to note that 
$V_{A}(\vec r)$ and $V_{B}(\vec r)$ have opposite signs
and 
their combined effect 
$V_{AB}(\vec r) = V_{A}(\vec r)+ V_{B}(\vec r)$
depends on their relative charges.
Using the charge neutrality $q_B=3q_o-q_A$,
we can write 
\begin{align}
V_{AB}(\vec r) &= f_{q_A}V_{O}(\vec r),\\
f_{q_A} &= \left(\frac{1}{32}+\frac{8}{81 \sqrt{3}}\right)\frac{q_A}{q_o} -\frac{3}{32},
\end{align}
which shows that ${V_{AB}(\vec r)}$ vanishes
at
$q_A/q_o=729/({256 \sqrt{3}+243}) {\simeq 1.062}$.
%where they cancel each other exactly.

@@

Before we present our results,
let us briefly describe 
the kind of behaviour that would indicate a JT instability 
against the tilting and rotations of the octahedra,
and how it relates to the standard JT problem.
As describe in Sec.~\ref{sec:JT},
the optimum structure can have a finite distortion
as long as 
the splitting in the single particle spectrum has a different dependence on the distortion
than the elastic energy associated with the changes in the unit cell volume.
In contrast to 
%hypothetical 
elongations or 
compressions---$a\delta_1,a\delta_2$ 
of an octahedron along two of its axes---where the 
 the unit cell volume changes
as $\sim a^3\left(\delta_1+\delta_1\right)$
and thus the elastic energy approximately goes
as $\sim\left(\delta_1+\delta_1\right)^2$,
the unit cell volume
in case of octahedral rotation and tilting
 depends on the ``distortions'' $\alpha,\beta$ 
as ${v=4a^3 \cos ^2\alpha \cos ^2\beta}$,
and, at $\alpha,\beta\ll1$,
it changes as $\sim a^3\left(\alpha ^2+\beta ^2\right)$
leading to a \emph{quartic} elastic energy that should roughly behave as
$\sim\left(\alpha ^2+\beta ^2\right)^2$.
So, the behaviour we are looking for in the electronic spectrum 
is not linear
but quadratic, 
which would compete with such a quartic elastic energy
 to produce a finite distortion in the optimum structure.
Thinking of squared angles 
$\alpha ^2,\beta ^2$
as the basic distortion variables,
the order of the electronic and elastic energy terms
matches to the standard JT problem.

@@

\begin{figure}[htbp]
\begin{center}
%\includegraphics[width=1.0\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/full-lattice-potential-tests/full-lattice-vs-AB}
\includegraphics[width=1.0\linewidth]{full-lattice-vs-AB}
\caption{
The effect of the potential of full ionic lattice.
$t_{2g}$ spectrum computed using the potential of full lattice 
and that obtained from only nearest AB cages (and the anion octahedron itself)
are
shown as a function of $q_A/q_o$ at 
$\alpha=0,\beta=20^\circ$, $\Delta_o=4q_o/15a$.
Full lattice only shifts the balance between the two cation cages towards A 
(i.e, the crossing point, where the two cages balance each other, moves to smaller $q_A$)
but keeps the qualitative picture the same.
}
\label{fig:t2g-lattice}
\end{center}
\end{figure}

@@

\label{Bibliography}
\bibliographystyle{apsrev4-1}
%\bibliography{perovskites}

%merlin.mbs apsrev4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
%Control: key (0)
%Control: author (8) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (-1) disabled
%Control: page (0) single
%Control: year (1) truncated
%Control: production of eprint (0) enabled
\begin{thebibliography}{83}%
\makeatletter
\providecommand \@ifxundefined [1]{%
 \@ifx{#1\undefined}
}%
\providecommand \@ifnum [1]{%
 \ifnum #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \@ifx [1]{%
 \ifx #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \natexlab [1]{#1}%
\providecommand \enquote  [1]{``#1''}%
\providecommand \bibnamefont  [1]{#1}%
\providecommand \bibfnamefont [1]{#1}%
\providecommand \citenamefont [1]{#1}%
\providecommand \href@noop [0]{\@secondoftwo}%
\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
\providecommand \@href[1]{\@@startlink{#1}\@@href}%
\providecommand \@@href[1]{\endgroup#1\@@endlink}%
\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
\providecommand \@@startlink[1]{}%
\providecommand \@@endlink[0]{}%
\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
\providecommand \urlprefix  [0]{URL }%
\providecommand \Eprint [0]{\href }%
\providecommand \doibase [0]{http://dx.doi.org/}%
\providecommand \selectlanguage [0]{\@gobble}%
\providecommand \bibinfo  [0]{\@secondoftwo}%
\providecommand \bibfield  [0]{\@secondoftwo}%
\providecommand \translation [1]{[#1]}%
\providecommand \BibitemOpen [0]{}%
\providecommand \bibitemStop [0]{}%
\providecommand \bibitemNoStop [0]{.\EOS\space}%
\providecommand \EOS [0]{\spacefactor3000\relax}%
\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
\let\auto@bib@innerbib\@empty
%</preamble>
\bibitem [{\citenamefont {Zubko}\ \emph {et~al.}(2011)\citenamefont {Zubko},
  \citenamefont {Gariglio}, \citenamefont {Gabay}, \citenamefont {Ghosez},\
  and\ \citenamefont {Triscone}}]{ZubkoARCMP11}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
  {Zubko}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Gariglio}},
  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Gabay}}, \bibinfo
  {author} {\bibfnamefont {P.}~\bibnamefont {Ghosez}}, \ and\ \bibinfo {author}
  {\bibfnamefont {J.-M.}\ \bibnamefont {Triscone}},\ }\href {\doibase
  10.1146/annurev-conmatphys-062910-140445} {\bibfield  {journal} {\bibinfo
  {journal} {Annual Review of Condensed Matter Physics}\ }\textbf {\bibinfo
  {volume} {2}},\ \bibinfo {pages} {141} (\bibinfo {year} {2011})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Hwang}\ \emph {et~al.}(2012)\citenamefont {Hwang},
  \citenamefont {Iwasa}, \citenamefont {Kawasaki}, \citenamefont {Keimer},
  \citenamefont {Nagaosa},\ and\ \citenamefont {Tokura}}]{HwangNM12}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.~Y.}\ \bibnamefont
  {Hwang}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Iwasa}},
  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Kawasaki}}, \bibinfo
  {author} {\bibfnamefont {B.}~\bibnamefont {Keimer}}, \bibinfo {author}
  {\bibfnamefont {N.}~\bibnamefont {Nagaosa}}, \ and\ \bibinfo {author}
  {\bibfnamefont {Y.}~\bibnamefont {Tokura}},\ }\href {\doibase
  10.1038/nmat3223} {\bibfield  {journal} {\bibinfo  {journal} {Nature
  Materials}\ }\textbf {\bibinfo {volume} {11}},\ \bibinfo {pages} {103}
  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Bhattacharya}\ and\ \citenamefont
  {May}(2014)}]{BhattacharyaARMR14}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Bhattacharya}}\ and\ \bibinfo {author} {\bibfnamefont {S.~J.}\ \bibnamefont
  {May}},\ }\href {\doibase 10.1146/annurev-matsci-070813-113447} {\bibfield
  {journal} {\bibinfo  {journal} {Annual Review of Materials Research}\
  }\textbf {\bibinfo {volume} {44}},\ \bibinfo {pages} {65} (\bibinfo {year}
  {2014})},\ \Eprint
  {http://arxiv.org/abs/https://doi.org/10.1146/annurev-matsci-070813-113447}
  {https://doi.org/10.1146/annurev-matsci-070813-113447} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {Hellman}\ \emph {et~al.}(2017)\citenamefont
  {Hellman}, \citenamefont {Hoffmann}, \citenamefont {Tserkovnyak},
  \citenamefont {Beach}, \citenamefont {Fullerton}, \citenamefont {Leighton},
  \citenamefont {MacDonald}, \citenamefont {Ralph}, \citenamefont {Arena},
  \citenamefont {D\"urr}, \citenamefont {Fischer}, \citenamefont {Grollier},
  \citenamefont {Heremans}, \citenamefont {Jungwirth}, \citenamefont {Kimel},
  \citenamefont {Koopmans}, \citenamefont {Krivorotov}, \citenamefont {May},
  \citenamefont {Petford-Long}, \citenamefont {Rondinelli}, \citenamefont
  {Samarth}, \citenamefont {Schuller}, \citenamefont {Slavin}, \citenamefont
  {Stiles}, \citenamefont {Tchernyshyov}, \citenamefont {Thiaville},\ and\
  \citenamefont {Zink}}]{HellmanRMP17}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
  {Hellman}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Hoffmann}},
  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Tserkovnyak}}, \bibinfo
  {author} {\bibfnamefont {G.~S.~D.}\ \bibnamefont {Beach}}, \bibinfo {author}
  {\bibfnamefont {E.~E.}\ \bibnamefont {Fullerton}}, \bibinfo {author}
  {\bibfnamefont {C.}~\bibnamefont {Leighton}}, \bibinfo {author}
  {\bibfnamefont {A.~H.}\ \bibnamefont {MacDonald}}, \bibinfo {author}
  {\bibfnamefont {D.~C.}\ \bibnamefont {Ralph}}, \bibinfo {author}
  {\bibfnamefont {D.~A.}\ \bibnamefont {Arena}}, \bibinfo {author}
  {\bibfnamefont {H.~A.}\ \bibnamefont {D\"urr}}, \bibinfo {author}
  {\bibfnamefont {P.}~\bibnamefont {Fischer}}, \bibinfo {author} {\bibfnamefont
  {J.}~\bibnamefont {Grollier}}, \bibinfo {author} {\bibfnamefont {J.~P.}\
  \bibnamefont {Heremans}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
  {Jungwirth}}, \bibinfo {author} {\bibfnamefont {A.~V.}\ \bibnamefont
  {Kimel}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Koopmans}},
  \bibinfo {author} {\bibfnamefont {I.~N.}\ \bibnamefont {Krivorotov}},
  \bibinfo {author} {\bibfnamefont {S.~J.}\ \bibnamefont {May}}, \bibinfo
  {author} {\bibfnamefont {A.~K.}\ \bibnamefont {Petford-Long}}, \bibinfo
  {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Rondinelli}}, \bibinfo
  {author} {\bibfnamefont {N.}~\bibnamefont {Samarth}}, \bibinfo {author}
  {\bibfnamefont {I.~K.}\ \bibnamefont {Schuller}}, \bibinfo {author}
  {\bibfnamefont {A.~N.}\ \bibnamefont {Slavin}}, \bibinfo {author}
  {\bibfnamefont {M.~D.}\ \bibnamefont {Stiles}}, \bibinfo {author}
  {\bibfnamefont {O.}~\bibnamefont {Tchernyshyov}}, \bibinfo {author}
  {\bibfnamefont {A.}~\bibnamefont {Thiaville}}, \ and\ \bibinfo {author}
  {\bibfnamefont {B.~L.}\ \bibnamefont {Zink}},\ }\href {\doibase
  10.1103/RevModPhys.89.025006} {\bibfield  {journal} {\bibinfo  {journal}
  {Rev. Mod. Phys.}\ }\textbf {\bibinfo {volume} {89}},\ \bibinfo {pages}
  {025006} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Chen}\ and\ \citenamefont
  {Millis}(2017)}]{ChenJPCM17}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {Chen}}\ and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Millis}},\
  }\href {\doibase 10.1088/1361-648X/aa6efe} {\bibfield  {journal} {\bibinfo
  {journal} {Journal of Physics: Condensed Matter}\ }\textbf {\bibinfo {volume}
  {29}},\ \bibinfo {pages} {243001} (\bibinfo {year} {2017})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Dagotto}(2005)}]{DagottoS05}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
  {Dagotto}},\ }\href {\doibase 10.1126/science.1107559} {\bibfield  {journal}
  {\bibinfo  {journal} {Science}\ }\textbf {\bibinfo {volume} {309}},\ \bibinfo
  {pages} {257} (\bibinfo {year} {2005})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Dagotto}\ and\ \citenamefont
  {Tokura}(2008)}]{DagottoMRSB08}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
  {Dagotto}}\ and\ \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
  {Tokura}},\ }\href {\doibase 10.1557/mrs2008.223} {\bibfield  {journal}
  {\bibinfo  {journal} {MRS Bulletin}\ }\textbf {\bibinfo {volume} {33}},\
  \bibinfo {pages} {1037} (\bibinfo {year} {2008})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Salamon}\ and\ \citenamefont
  {Jaime}(2001)}]{SalamonRMP01}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~B.}\ \bibnamefont
  {Salamon}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Jaime}},\ }\href {\doibase 10.1103/RevModPhys.73.583} {\bibfield  {journal}
  {\bibinfo  {journal} {Rev. Mod. Phys.}\ }\textbf {\bibinfo {volume} {73}},\
  \bibinfo {pages} {583} (\bibinfo {year} {2001})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Wang}\ \emph {et~al.}(2003)\citenamefont {Wang},
  \citenamefont {Neaton}, \citenamefont {Zheng}, \citenamefont {Nagarajan},
  \citenamefont {Ogale}, \citenamefont {Liu}, \citenamefont {Viehland},
  \citenamefont {Vaithyanathan}, \citenamefont {Schlom}, \citenamefont
  {Waghmare}, \citenamefont {Spaldin}, \citenamefont {Rabe}, \citenamefont
  {Wuttig},\ and\ \citenamefont {Ramesh}}]{WangS03}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {Wang}}, \bibinfo {author} {\bibfnamefont {J.~B.}\ \bibnamefont {Neaton}},
  \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Zheng}}, \bibinfo
  {author} {\bibfnamefont {V.}~\bibnamefont {Nagarajan}}, \bibinfo {author}
  {\bibfnamefont {S.~B.}\ \bibnamefont {Ogale}}, \bibinfo {author}
  {\bibfnamefont {B.}~\bibnamefont {Liu}}, \bibinfo {author} {\bibfnamefont
  {D.}~\bibnamefont {Viehland}}, \bibinfo {author} {\bibfnamefont
  {V.}~\bibnamefont {Vaithyanathan}}, \bibinfo {author} {\bibfnamefont {D.~G.}\
  \bibnamefont {Schlom}}, \bibinfo {author} {\bibfnamefont {U.~V.}\
  \bibnamefont {Waghmare}}, \bibinfo {author} {\bibfnamefont {N.~A.}\
  \bibnamefont {Spaldin}}, \bibinfo {author} {\bibfnamefont {K.~M.}\
  \bibnamefont {Rabe}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Wuttig}}, \ and\ \bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
  {Ramesh}},\ }\href {\doibase 10.1126/science.1080615} {\bibfield  {journal}
  {\bibinfo  {journal} {Science}\ }\textbf {\bibinfo {volume} {299}},\ \bibinfo
  {pages} {1719} (\bibinfo {year} {2003})},\ \Eprint
  {http://arxiv.org/abs/https://www.science.org/doi/pdf/10.1126/science.1080615}
  {https://www.science.org/doi/pdf/10.1126/science.1080615} \BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Dawber}\ \emph {et~al.}(2005)\citenamefont {Dawber},
  \citenamefont {Rabe},\ and\ \citenamefont {Scott}}]{DawberRMP05}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
  {Dawber}}, \bibinfo {author} {\bibfnamefont {K.~M.}\ \bibnamefont {Rabe}}, \
  and\ \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Scott}},\ }\href
  {\doibase 10.1103/RevModPhys.77.1083} {\bibfield  {journal} {\bibinfo
  {journal} {Rev. Mod. Phys.}\ }\textbf {\bibinfo {volume} {77}},\ \bibinfo
  {pages} {1083} (\bibinfo {year} {2005})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Schlom}\ \emph {et~al.}(2007)\citenamefont {Schlom},
  \citenamefont {Chen}, \citenamefont {Eom}, \citenamefont {Rabe},
  \citenamefont {Streiffer},\ and\ \citenamefont {Triscone}}]{SchlomARMR07}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.~G.}\ \bibnamefont
  {Schlom}}, \bibinfo {author} {\bibfnamefont {L.-Q.}\ \bibnamefont {Chen}},
  \bibinfo {author} {\bibfnamefont {C.-B.}\ \bibnamefont {Eom}}, \bibinfo
  {author} {\bibfnamefont {K.~M.}\ \bibnamefont {Rabe}}, \bibinfo {author}
  {\bibfnamefont {S.~K.}\ \bibnamefont {Streiffer}}, \ and\ \bibinfo {author}
  {\bibfnamefont {J.-M.}\ \bibnamefont {Triscone}},\ }\href {\doibase
  10.1146/annurev.matsci.37.061206.113016} {\bibfield  {journal} {\bibinfo
  {journal} {Annual Review of Materials Research}\ }\textbf {\bibinfo {volume}
  {37}},\ \bibinfo {pages} {589} (\bibinfo {year} {2007})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Koster}\ \emph {et~al.}(2012)\citenamefont {Koster},
  \citenamefont {Klein}, \citenamefont {Siemons}, \citenamefont {Rijnders},
  \citenamefont {Dodge}, \citenamefont {Eom}, \citenamefont {Blank},\ and\
  \citenamefont {Beasley}}]{KosterRMP12}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
  {Koster}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Klein}},
  \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Siemons}}, \bibinfo
  {author} {\bibfnamefont {G.}~\bibnamefont {Rijnders}}, \bibinfo {author}
  {\bibfnamefont {J.~S.}\ \bibnamefont {Dodge}}, \bibinfo {author}
  {\bibfnamefont {C.-B.}\ \bibnamefont {Eom}}, \bibinfo {author} {\bibfnamefont
  {D.~H.~A.}\ \bibnamefont {Blank}}, \ and\ \bibinfo {author} {\bibfnamefont
  {M.~R.}\ \bibnamefont {Beasley}},\ }\href {\doibase
  10.1103/RevModPhys.84.253} {\bibfield  {journal} {\bibinfo  {journal} {Rev.
  Mod. Phys.}\ }\textbf {\bibinfo {volume} {84}},\ \bibinfo {pages} {253}
  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Bersuker}(1975)}]{BersukerCCR75}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {I.}~\bibnamefont
  {Bersuker}},\ }\href {\doibase https://doi.org/10.1016/S0010-8545(00)80266-0}
  {\bibfield  {journal} {\bibinfo  {journal} {Coordination Chemistry Reviews}\
  }\textbf {\bibinfo {volume} {14}},\ \bibinfo {pages} {357} (\bibinfo {year}
  {1975})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Kugel}\ and\ \citenamefont {Khomski{\u
  i}}(1982)}]{KugelSPU82}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.~I.}\ \bibnamefont
  {Kugel}}\ and\ \bibinfo {author} {\bibfnamefont {D.~I.}\ \bibnamefont
  {Khomski{\u i}}},\ }\href {\doibase 10.1070/PU1982v025n04ABEH004537}
  {\bibfield  {journal} {\bibinfo  {journal} {Soviet Physics Uspekhi}\ }\textbf
  {\bibinfo {volume} {25}},\ \bibinfo {pages} {231} (\bibinfo {year}
  {1982})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Millis}(1998)}]{MillisN98}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~J.}\ \bibnamefont
  {Millis}},\ }\href {\doibase 10.1038/32348} {\bibfield  {journal} {\bibinfo
  {journal} {Nature}\ }\textbf {\bibinfo {volume} {392}},\ \bibinfo {pages}
  {147} (\bibinfo {year} {1998})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Terakura}(2007)}]{TerakuraPMS07}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
  {Terakura}},\ }\href {\doibase https://doi.org/10.1016/j.pmatsci.2006.10.007}
  {\bibfield  {journal} {\bibinfo  {journal} {Progress in Materials Science}\
  }\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {388} (\bibinfo {year}
  {2007})},\ \bibinfo {note} {modelling electrons and atoms for materials
  science}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Halcrow}(2013)}]{HalcrowRSC13}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~A.}\ \bibnamefont
  {Halcrow}},\ }\href {\doibase 10.1039/C2CS35253B} {\bibfield  {journal}
  {\bibinfo  {journal} {Chemical Society Reviews}\ }\textbf {\bibinfo {volume}
  {42}},\ \bibinfo {pages} {1784} (\bibinfo {year} {2013})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Lufaso}\ and\ \citenamefont
  {Woodward}(2004)}]{LufasoAC04}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.~W.}\ \bibnamefont
  {Lufaso}}\ and\ \bibinfo {author} {\bibfnamefont {P.~M.}\ \bibnamefont
  {Woodward}},\ }\href {\doibase 10.1107/S0108768103026661} {\bibfield
  {journal} {\bibinfo  {journal} {Acta Crystallographica Section B}\ }\textbf
  {\bibinfo {volume} {60}},\ \bibinfo {pages} {10} (\bibinfo {year}
  {2004})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Glazer}(1972)}]{GlazerAC72}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {Glazer}},\ }\href {\doibase 10.1107/S0567740872007976} {\bibfield  {journal}
  {\bibinfo  {journal} {Acta Crystallogr., Sect. B: Struct. Crystallogr. Cryst.
  Chem.}\ }\textbf {\bibinfo {volume} {28}},\ \bibinfo {pages} {3384} (\bibinfo
  {year} {1972})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Thomas}(1996)}]{ThomasAC96}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~W.}\ \bibnamefont
  {Thomas}},\ }\href {\doibase 10.1107/S0108768195006100} {\bibfield  {journal}
  {\bibinfo  {journal} {Acta Crystallographica Section B}\ }\textbf {\bibinfo
  {volume} {52}},\ \bibinfo {pages} {16} (\bibinfo {year} {1996})}\BibitemShut

>>>>

%\subsection{Parameters $D_2$ and $D_4$ of the JT ion/atom}
The terms in square brackets on the right of Eqs.~\ref{eq:JTt2g},\ref{eq:JTeg}
contain $D_2$ and $D_4$. 
$D_4$ is proportional to the crystal field splitting $\Delta$ 
($D_4=\frac{3a}{160q_o} \Delta$) and 
we can estimate its value from realistic values of $\Delta$ for a given octahedron of size $a$ and ion charge $-q_o$. 
Assuming $a=7.0$ bohr and $q_o=2e$ ($e$ being the elementary charge),
$D_4$ ranges from $1 \times 10^{-3}$ to  $7 \times 10^{-3}$
for $\Delta \simeq 0.5-3$~eV.
Typically, ${D_{2}\approx \sqrt{D_{4}}}$,
so $D_{2}/D_{4} \approx 1/\sqrt{D_{4}}$
and it roughly ranges between $29$ and $12$ for $D_4$ in the above range.
This means that, for realistic systems, 
we should always have $D_{2}/D_{4} > 100/27, 25/9 \approx 4$
so 
$D_{2}$ terms in 
Eqs.~\ref{eq:JTt2g},\ref{eq:JTeg}
dominate and determine the sign and size of the splittings.

@@

In a cubic perovskite ABO$_3$,
each transition metal atom/ion B
is surrounded by six O$^{2-}$, each at a distance $a/2$,
 making an octahedron of size $a$ around it.
The potential of these ions $V_{O}(\vec r)= V_{oct}(\vec r)$
splits the degenerate d-orbitals
by an amount ${\Delta_o=160q_oD_{4}/3a}$
 into $t_{2g}$ and $e_g$.
However, this is not the only crystal field potential at the B site.
The coordination cages of next nearest neighbours, 
eight A atoms/cations all at the same distance $\sqrt{3}a/2$
make a cube around the B ion,
which also produces an octahedral field, given by 
\begin{align}
V_{A}(\vec r) = \frac{8}{81 \sqrt{3}}\frac{q_A}{q_o} V_{O}(\vec r).
\end{align}
Furthermore,
the nearest neighbouring B ions, six of them each at a distance $a$,
make an octahedron, with a potential 
\begin{align}
V_{B}(\vec r) = -\frac{1}{32}\frac{q_B}{q_o}V_{O}(\vec r).
\end{align}
It is interesting to note that 
$V_{A}(\vec r)$ and $V_{B}(\vec r)$ have opposite signs
and 
their combined effect 
$V_{AB}(\vec r) = V_{A}(\vec r)+ V_{B}(\vec r)$
depends on their relative charges.
Using the charge neutrality $q_B=3q_o-q_A$ (remember that we take anion charge to be $-q_o$),
we can write 
\begin{align}
V_{AB}(\vec r) &= f_{q_A}V_{O}(\vec r),\\
f_{q_A} &= \left(\frac{1}{32}+\frac{8}{81 \sqrt{3}}\right)\frac{q_A}{q_o} -\frac{3}{32},
\end{align}
which shows that ${V_{AB}(\vec r)}$ vanishes
at
$q_A/q_o=729/({256 \sqrt{3}+243}) {\simeq 1.062}$.
%where they cancel each other exactly.

@@

Before we present our results,
let us briefly describe 
the kind of behaviour that would indicate a JT instab against the tilting and rotations of the octahedra,
and how it relates to the standard JT problem.
As describe in Sec.~\ref{sec:JT},
the optimum structure can have a finite distortion
as long as 
the splitting in the single particle spectrum has a different dependence on the distortion
than the elastic energy associated with the changes in the unit cell volume.
\rev{Here, the unit cell volume
 depends on the ``distortions'' $\alpha,\beta$ 
as ${v=4a^3 \cos ^2\alpha \cos ^2\beta}$,
and, at $\alpha,\beta\ll1$,
it changes as $\sim a^3\left(\alpha ^2+\beta ^2\right)$.
The elastic energy in the harmonic approximation thus goes
as $\sim\left(\alpha ^2+\beta ^2\right)^2$,
which is \emph{quartic} in $\alpha,\beta$.
So, 
in contrast to the case of
elongations or 
compressions  
of the octahedra~\cite{KhomskiiCR21},
the behavior we are looking for in the electronic spectrum 
is not linear
but quadratic, 
which would compete with such a quartic elastic energy
 to produce a finite distortion in the optimum structure.
Thinking of squared angles 
$\alpha ^2,\beta ^2$
as the basic distortion variables,
the order of the electronic and elastic energy terms
matches to the standard JT problem.
}

\com{In contrast to 
%hypothetical 
elongations or 
compressions---$a\delta_1,a\delta_2$ 
of an octahedron along two of its axes---where the 
 the unit cell volume changes
as $\sim a^3\left(\delta_1+\delta_2\right)$
and thus
the elastic energy in the harmonic approximation goes
as $\sim\left(\delta_1+\delta_2\right)^2$,
the unit cell volume
in case of octahedral rotation and tilting
 depends on the ``distortions'' $\alpha,\beta$ 
as ${v=4a^3 \cos ^2\alpha \cos ^2\beta}$,
and, at $\alpha,\beta\ll1$,
it changes as $\sim a^3\left(\alpha ^2+\beta ^2\right)$
leading to a \emph{quartic} elastic energy that should roughly behave as
$\sim\left(\alpha ^2+\beta ^2\right)^2$.
So, the behavior we are looking for in the electronic spectrum 
is not linear
but quadratic, 
which would compete with such a quartic elastic energy
 to produce a finite distortion in the optimum structure.
Thinking of squared angles 
$\alpha ^2,\beta ^2$
as the basic distortion variables,
the order of the electronic and elastic energy terms
matches to the standard JT problem.
}

@@

\begin{figure}[htbp]
\begin{center}
%\includegraphics[width=1.0\linewidth]{/Users/maz/repos/perovskite/calc/analytical-structure-param/full-lattice-potential-tests/full-lattice-vs-AB}
\includegraphics[width=1.0\linewidth]{full-lattice-vs-AB}
\caption{
The effect of the potential of full ionic lattice.
$t_{2g}$ spectrum computed using the potential of full lattice 
and that obtained from only nearest AB cages (and the anion octahedron itself)
are
shown as a function of $q_A/q_o$ at 
$\alpha=0,\beta=20^\circ$, $\Delta_o=4q_o/15a$.
Full lattice only shifts the balance between the two cation cages towards A 
(i.e, the crossing point, where the two cages balance each other, moves to smaller $q_A$)
but keeps the qualitative picture the same.
(The bold black line is nothing but an overlap of two practically degenerate black lines.)
}
\label{fig:t2g-lattice}
\end{center}
\end{figure}

@@

\label{Bibliography}
%\bibliographystyle{apsrev4-1}
\bibliography{perovskites}
